---
title: "FA2 (PROBABILITY & PROBABILITY DISTRIBUTIONS"
author: "QUINTERO"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Use R to illustrate that the probability of getting:
(a) a head is 0.5 if a fair coin is tossed repeatedly;

```{r}
Coinflip = function(x) sample(0:1,x,rep=T)
total=Coinflip(10)
total

#estimation (TAILS)
sum (total == 0)/10
#estimation (HEADS)
sum (total == 1)/10
```

(b) a red card is 0.5 if cards are drawn repeatedly with replacement from awell-shuffled deck;

```{r}
Cardflip = function(x) sample(1:52,x,rep=T)
total=Cardflip(52)
total

#estimation (RED)
sum (total >= 26)/52
#estimation (BLACK)
sum (total < 26)/52

```

(c) an even number is 0.5 if a fair die is rolled repeatedly.

```{r}
sample(1:6,6,rep=T)
dieroll = function(x) sample(1:6,x,rep=T)
total=dieroll(6)
total

#estimation (EVEN)
sum1 <- sum(total == 2)/6
sum2 <-  sum(total == 4)/6
sum3 <- sum(total == 6)/6
sum4 <- sum1 + sum2 + sum3
sum4

#estimation (ODD)
sum5 <- sum(total == 1)/6
sum6 <- sum(total == 3)/6
sum7 <- sum(total == 5)/6
sum8 <- sum5 + sum6 + sum7
sum8

```

## 2. An experiment consists of tossing two fair coins. Use R to simulate this experiment 100 times and obtain the relative frequency of each possible outcome. Hence, estimate the probability of getting one head and one tail in any order.

```{r}
Coinflip = function(x) sample(0:1,x,rep=T)
total=Coinflip(100)
total

#estimation (TAILS)
sum (total == 0)/100
#estimation (HEADS)
sum (total == 1)/100
```
