---
title: 'FA 3: PROBABILITY AND PROBABILITY DISTRIBUTION'
author: "QUINTERO"
date: "2024-02-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.  A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, determine the probability that:

# A.  a 1 was received;

```{r}
# Given probabilities
prob_R0_given_T0 <- 0.95       
prob_R1_given_T1 <- 0.75       
prob_T0 <- 0.70                

# Calculate the complement
prob_R1_given_T1_comp <- 1 - prob_T0 * (1 - prob_R1_given_T1)

# (a) Probability that a 1 was received
prob_R1 <- prob_T0 * (1 - prob_R0_given_T0) + (1 - prob_T0) * prob_R1_given_T1_comp

# (b) Probability that a 1 was transmitted given that a 1 was received
prob_T1_given_R1 <- prob_T0 * prob_R0_given_T0 / prob_R1

# Print the results
cat("Probability that a 1 was received:", prob_R1, "\n")
```

# B. a 1 was transmitted given than a 1 was received. 

```{r}
cat("Probability that a 1 was transmitted given that a 1 was received:", prob_T1_given_R1, "\n")
```

# 2. There are three employees working at an IT company: Jane, Amy, and Ava, doing 10%, 30%, and 60% of the programming, respectively. 8% of Jane’s work, 5% of Amy’s work, and just 1% of Ava‘s work is in error. What is the overall percentage of error? If a program is found with an error, who is the most likely person to have written it? 

```{r}
# Employee percentages
percentage <- c(0.10, 0.30, 0.60)

# Error rates 
error <- c(0.08, 0.05, 0.01)

# Employee names 
employees <- c("Jane", "Amy", "Ava")

# Calculate overall error
overallerror <- sum(percentage * error)

# Print the overall error rate
cat("Overall Error Rate:", overallerror, "\n")

# Identify the most likely person 
most_likely_person <- which.max(percentage * error)
most_likely_person_name <- employees[most_likely_person]

# Print the most likely employee name to have a error in her program
cat("Most likely employee to have a error in her program is:", most_likely_person_name, "\n")
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
